---
title: "Stage 1 Sample Design with replacements"
subtitle: "by Will MacKenzie, Kiri Daust, Gen Perkins"
Methods author: "Will MacKenzie"
Script authors: "Kiri Daust; Will MacKenzie; Colin Chisholm; Gen Perkins, Matt Coghill" 
date: "09/08/2022"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r global_options, include=FALSE}
require(knitr)
```

# Sample Design with Replacement

After the creation of stage 1 sample plan, there are occasions when a single or multiple sites need to be changed or reselected. In this case there are multiple methods; 1) Manual replacement or 2) automated replacement. 

in both cases replacement is based on finding equivalent environmental space for the site which has been rejected. The script below used base layers created in the development of sample plan script (03a_Stage1_sampleDesign.Rmd) to replace sites while accounting for sites which are already layed out. 

This script requires: 

- exising s1_sampling.gpks
- covarates used in sample design (i.e 25m; DAH, landscape, etc) or the Landscape Validation layer, created in 02a_DEM_SpatialLayer_Prep.R script. 


```{r setup, include=FALSE}
rm(list = ls())


#library(velox)
library(raster)
library(sf)
library(fasterize)
library(gdistance)
library(tidyverse)
library(LearnGeom)
library(stars)
library(plyr)
library(data.table)
library(reticulate)
#devtools::install_github ("pierreroudier/clhs") 
library(clhs)
library(foreach)
library(sp)
library(foreach)
library(here)
library(stats)
library(raster)
library(tidyverse)
library(plyr)

```


# Option 1: manually replace a single site (stand alone option)

```{r}

# Date Creek replacement sites 
# need to sample
 badPointNames <- c('ICH mc 2_2.2_7_cLHS',
                   'ICH mc 2_2.5_10',
                   'ICH mc 2_3.2_12',
                   'ICH mc 2_3.5_15',
                   'ICH mc 2_4.4_18',
                   'ICH mc 2_5.2_22',
                   'ICH mc 2_5.4_24_cLHS')

aoi <- "DateCreek"
sampleplan <- "s1_sampling_V3.gpkg"

transect_folder <- file.path(paste0(aoi, "_AOI"), "2_sample_design", "stage1_StudyDesign",
                             "transect_layout")
raster_folder <- file.path(paste0(aoi, "_AOI"), "1_map_inputs", "covariates",
                           "25m")

st_layers(file.path(transect_folder, sampleplan))

points <- st_read(file.path(transect_folder, sampleplan ), "s1_sampling_V3")

covariates <- list.files(raster_folder,full.names = T)
covariates <- covariates[grep("tif",covariates)]
covariates <- covariates[c(2,3,9,12)]

ancDat <- raster::stack(covariates)

badPoints <- points[points$id %in% badPointNames, c("id","geom")]

new_sample_space <- function(badPoints,ancDat,basename){##
  for(i in 1:nrow(badPoints)){
    cat("Processing point",i)
    currPt <- badPoints[i,]
    covarVals <- extract(ancDat,currPt)
    sArea <- calc(ancDat, fun = function(x){if(x[1] == covarVals[1,1] 
                                               & x[2] == covarVals[1,2] 
                                               & x[3] == covarVals[1,3] 
                                               & x[4] == covarVals[1,4]) 1 else 0})
    writeRaster(sArea, file.path(transect_folder, paste0("NewSampleArea_",currPt$id,".tif")),format = "GTiff", overwrite = TRUE)
  }
  return(TRUE)
}

new_sample_space(badPoints,ancDat,"NewSpace_ICHmc2")
```



# Option 2: Replace sites using entire sample design 

This option will re-create missing sites within the slices of the entire S1 sample Design. 
Incorporating previous sampling plots or adjusting the slice if sites are unavailable. 


```{r}
# input parameters

#AOI <- "Deception"
AOI <- "DateCreek"

map_res <- 25 # Defines resolution of map to use for cLHS
transect_length <- 250 # Define the length of triangle on each side
heli = TRUE

map_res_character <- paste0(map_res, "m")
AOI.dir <- file.path(paste0(AOI, "_AOI"))
shapes_path <- file.path(AOI.dir, "0_raw_inputs", "base_layers")
raster_path <- file.path(AOI.dir, "1_map_inputs", "covariates", map_res_character)
out_path <- file.path(AOI.dir, "2_sample_design", "stage1_StudyDesign")
sampling_raw_folder <- file.path(out_path, "input_raster")
clhs_outpath <- file.path(out_path, "clhs_sample_plans")
mask_path <- file.path(AOI.dir, "2_sample_design", "stage1_StudyDesign","input_raster")

centroid_distance <- 400
rad_exclusion <- 1000

```


```{r}

# read in covars
covars <- paste(raster_path, c("DAH_LS.tif","Landform_LSs.tif","MRVBF_LSs.tif"), sep = "/")
layerNames <- c("DAH","LFC","MRVBF","cost") ##need to change this if you change the layers
ancDat <- raster::stack(covars)
proj4string(ancDat) <- "+init=epsg:3005"


# read in slope data
slope <- raster(file.path(raster_path, "slope.tif"))
proj4string(slope) <- "+init=epsg:3005"

# read dem
alt <- raster(paste0(raster_path, "/dem.tif"))
proj4string(alt) <- "+init=epsg:3005"

# read in sites that were previously sampled.  
transect_out_path <- file.path(out_path, "transect_layout")
prev_samples <- list.files(file.path(transect_out_path, "raw"), pattern = ".gpkg$")

# date Creek
prev_samples <- grep("previously_sampled_ESSFwv.gpkg", prev_samples, perl = TRUE, value = TRUE)

previous_sites  <- foreach(i = prev_samples, .combine = rbind) %do% {
      st_read(file.path(transect_out_path, "raw", i))
    }       

# # Deception verion 
# prev_samples <- grep("previous_*", prev_samples, perl = TRUE, value = TRUE)
# xx <- st_read(file.path(transect_out_path, "raw", prev_samples)) %>%
#   mutate(idval = transect_id) %>%
#   separate(idval, c("type", "bgc","slice_num","slice_size"), "_") %>%
#   select(-c(type, bgc)) %>%
#   mutate(total = seq(1,length(slice_size),1),
#           subzone = "ESSFmcw") %>%
#   rowwise %>%
#   mutate(slice_num = as.numeric(slice_num) + 1)
# 
# previous_sites  <- xx


# for each of the bec units 
  
boi <- "ESSFwv"

#for (b in boi) {
 
  b <- boi[1]
    
  if (heli == TRUE) {
    
    ##road and lake buffer
    buff <- st_read(file.path(mask_path, paste0(b, "_exclude_poly.gpkg"))) %>% 
      st_transform(3005)
    
    ## clip to just ESSF
    boundary <- st_read(paste0(shapes_path ,"/bec.gpkg"))
    boundary <- boundary[,"BGC_LABEL"]
    boundary <- boundary[boundary$BGC_LABEL == b,] ## set as mask for individual BGC
    b2 <- st_union(boundary)
    b2 <- fasterize(boundary, slope)
    buff <- fasterize(buff, slope)
    buff <- mask(buff,b2)
    ancDat <- mask(ancDat,buff)
        
    ##read in drop points
    heliDrop <- st_read(paste0(shapes_path,"/heli_start_point2.gpkg")) %>%
      mutate(name = b)
    heliDrop <- heliDrop[,"name"]
    start <- as(heliDrop, "Spatial")

    
    # might need to switch this out for updated function 
    altDiff <- function(x){x[1]-x[2]}
    tr <- transition(alt, transitionFunction = altDiff, directions = 8, symm = F)
    tr <- geoCorrection(tr)
    adj <- adjacent(alt, cells = 1:ncell(alt), pairs = T, directions = 8)
    tr[adj] <- 1.5*exp(-3.5*abs(tr[adj] + 0.1)) ##tobler's hiking function
    tr <- geoCorrection(tr)

    acost <- accCost(tr, start)
    #acost <- acost/3600
    lays <- stack(ancDat,acost)
    names(lays) <- layerNames
    
    boi_mask <- raster(file.path(out_path, "input_raster", paste0(b,"_exclude_mask.tif"))) 
    names(boi_mask) = "delete"
    sample_layers_masked <- stack(lays, boi_mask) %>%  # add cost surface to stack
      raster::mask(boi_mask) 
    
    sample_layers_masked <- dropLayer(sample_layers_masked, 5)
    
  } else {
    
  # filter area able to place points
  boi_mask <- raster(file.path(out_path, "input_raster", paste0(b,"_exclude_mask.tif"))) 
  names(boi_mask) = "cost"
  sample_layers_masked <- stack(ancDat, boi_mask) %>%  # add cost surface to stack
    raster::mask(boi_mask) # mask by buffered cost surface
  names(sample_layers_masked) <- layerNames
    
  } # end of Heli or nor heli 
  
  for(rot in 2:10){
     # rot = 1
  
    # generate clhs points
      paired_distance <- 100
      num_samples <- 15
      slice_size <- 5
      num_slices <- round(num_samples/slice_size)
    
      # create a empty sf object and add points
      tmp <- st_sfc()
      class(tmp)[1] <- "sfc_POINT" # for points
      sample_points <- st_sf(DAH=integer(0),
                     MRVBF=integer(0),
                     LFC = integer(0),
                     cost=integer(0),
                     slice_num = integer(0), geometry=tmp)%>%
      st_set_crs(3005)

    # Index the previous sites to add to clhs. TO do this we need to 
      # convert the sp points and the raster brick to a data frame. 
      
    # convert raster brick to data frame and remove NA values
      lays_df <- as.data.frame(sample_layers_masked, xy = TRUE)
      lays_df <- lays_df  %>% drop_na()
      # convert the base layer to sf object 
      lays_sf <- st_as_sf(lays_df,coords = c("x","y")) %>%
        st_set_crs(3005)
      
      prev_xy <- previous_sites %>% filter(subzone == b)
      prev_xy <- cbind(prev_xy, st_coordinates(prev_xy)) %>%
          rename_with(tolower)
      
    #collect reference locations of lays data frame for previous samples.
     lays_prev <- lays_sf %>%
       mutate(rowno = as.integer(rownames(lays_sf))) %>%
       st_join(prev_xy) %>%
       filter(!is.na(id))

      # convert gemetry to match the data merge (convert geom to geometry)
      sample_points_all_slice <- prev_xy %>%
        st_drop_geometry() %>%
        st_as_sf(., coords = c("x","y")) %>%
        st_set_crs(3005)
      
      
      #buffer previous sites and asign a very high cost 
      
     #prev_site_cost_buff <- st_buffer(sample_points_all_slice, dist = rad_exclusion) 
      #             
      # # Check if points overlap buffer and then remove 
      # lays_df_SF <- lays_df %>%
      #      st_as_sf(., coords = c("x","y")) %>%
      #      st_set_crs(3005)
      #           
      # to_cost <- st_intersection(lays_df_SF, prev_site_cost_buff) %>%
      #       dplyr::select(geometry) %>%
      #       mutate(cost_update = 100000) 
      #             
      # # this part is slow and could be sped up..
      # lays_df_buffered <- st_join(lays_df_SF, to_cost) 
      #   
      # lays_df_buffered <- lays_df_buffered %>%
      #   mutate(cost = ifelse(!is.na(cost_update), cost_update, cost)) %>%
      #   select(-cost_update) 
      #             
      # lays_df <- cbind(st_coordinates(lays_df_buffered), lays_df_buffered) %>%
      #     st_drop_geometry() %>%
      #     dplyr::rename(x = X, y = Y)
      #             
     # for(i in 1:2){ 
     for(i in 1:num_slices){ # For each slice, perform cLHS (if there is sampleable area left from previous slices)
        print (paste0("slice: ",i))
         # i <- 1 #slice no
 
            # filter the sites already in the slice and check if extras needed
            prev_sample_points <- sample_points_all_slice %>% filter(slice_num == i)
            sites_required <- 5 - length(unique(prev_sample_points$id))
            lays_prev_slice <- lays_prev %>%  filter(slice_num == i)
                
            # check how many sites need to be genertad - if not rbind to data 
          if(sites_required > 0 ){
                   
            clhs_slice <- clhs(lays_sf,
                                size = slice_size,
                                must.include = as.matrix(c(lays_prev_slice$rowno)), 
                                iter = 10000,
                                simple = FALSE,
                                progress = TRUE,
                                cost= "cost",
                                use.cpp = T) # Run cLHS on the sampleable area
            
           clhs_sampled <- st_as_sf(clhs_slice$sampled_data, coords = c("x","y")) %>%
                    mutate(slice_num = i) %>%
                    dplyr::select(-c(DAH, LFC, MRVBF)) %>%
                    st_set_crs(3005)
                     
     
         for(j in 1:nrow(clhs_sampled)){ # Filter the close together samples from the cLHS run
           
                if(!is.na(clhs_sampled[j, ])){
                    distances <- data.frame(distance = st_distance(clhs_sampled, clhs_sampled[j, ])) %>%  
                      rownames_to_column() %>%
                      mutate_all(as.numeric) %>%
                      dplyr::filter(distance > rad_exclusion | distance == 0)
                      clhs_sampled <- clhs_sampled[distances$rowname, ]
                                              }
                  }
                     
                  # if number of sites is less than selected add a interator and this clhs willl be re-run using the repeat loop below  
         if(length(clhs_sampled$slice_num) < slice_size){
          
              print("points selected are less than sizes less than selected, re-running clhs")
                  
              # if the slice was not complete then rerun               

              tmp <- st_sfc()
                class(tmp)[1] <- "sfc_POINT" # for points
                sample_points_extra <- st_sf(DAH=integer(0),
                                 MRVBF=integer(0),
                                 LFC = integer(0),
                                 cost=integer(0),
                                 slice_num = integer(0), geometry=tmp)%>%
                st_set_crs(3005)

              repeat{
                  clhs_slice <- clhs(lays_sf,
                                size = slice_size,
                                must.include = as.matrix(c(lays_prev_slice$rowno)),  
                                iter = 10000,
                                simple = FALSE,
                                progress = TRUE,
                                cost= "cost",
                                use.cpp = T) # Run cLHS on the sampleable area
         
              clhs_sampled <- st_as_sf(clhs_slice$sampled_data, coords = c("x","y")) %>%
                  mutate(slice_num = i) %>%
                   dplyr::select(-c(DAH, LFC, MRVBF)) %>%
                  st_set_crs(3005)
         
        for(j in 1:nrow(clhs_sampled)){ # Filter the close together samples from the cLHS run
                       
                if(!is.na(clhs_sampled[j, ])){
                    distances <- data.frame(distance = st_distance(clhs_sampled, clhs_sampled[j, ])) %>%  
                      rownames_to_column() %>%
                      mutate_all(as.numeric) %>%
                      dplyr::filter(distance > rad_exclusion | distance == 0)
                      clhs_sampled <- clhs_sampled[distances$rowname, ]
                                              }
                  }
  
        if(length(clhs_sampled$slice_num)== slice_size) {
                
          print("correct number of pts generated")
                #sample_points_extra <- sample_points_extra
                #st_crs(sample_points_extra) <- 3005
                
                sample_points_extra <- rbind(clhs_sampled, sample_points_extra)
                
                # buffer the sites for the next iteration
                
                print("buffering selected points from sampling area")                   # if correct number of samples is selected then continue to next slice 
             
                clhs_sampled_buff <- st_buffer(sample_points_extra, dist = rad_exclusion) 
                
                to_cost <- st_intersection(lays_sf, clhs_sampled_buff) %>%
                   dplyr::select(geometry) %>%
                   mutate(cost_update = 100000) 
                  
              # this part is slow and could be sped up..
                lays_df_buffered <- st_join(lays_sf, to_cost)  %>%
                  mutate(cost = ifelse(!is.na(cost_update), cost_update, cost)) %>%
                  dplyr::select(-cost_update) 
      
                lays_sf<- lays_df_buffered
                #cbind(st_coordinates(lays_df_buffered), lays_df_buffered) %>%
                #  st_drop_geometry() %>%
                #  dplyr::rename(x = X, y = Y)
                
                
                 #buffer previous sites and asign a very high cost 
                      
        break # stop the repeat clhs loop if corect number achieved
         }
      # if the number is still less than required repeat the above code until correct number is produced
      print(paste0(length(clhs_sampled$slice_num), " sites generated, rerunning clhs"))
   
       }
  
          sample_points <- bind_rows(sample_points, sample_points_extra )
      
      #  } # end of extra repeats to add

               } else { # if the correct number of sites produced then skip to buffer 
              
              print("buffering selected points from sampling area")                   # if correct number of samples is selected then continue to next slice 
              clhs_sampled_buff <- st_buffer(clhs_sampled, dist = rad_exclusion) 
                  
             to_cost <- st_intersection(lays_sf, clhs_sampled_buff ) %>%
                dplyr::select(geometry) %>%
                mutate(cost_update = 100000) %>%
               distinct()
               
      
      # this part is slow and could be sped up..
             lays_df_buffered <- st_join(lays_sf, to_cost)  %>%
                mutate(cost = ifelse(!is.na(cost_update), cost_update, cost)) %>%
                dplyr::select(-cost_update) 
      
             lays_sf <- lays_df_buffered
      
              # add the sample points to the cumulative list
              sample_points <- bind_rows(sample_points, clhs_sampled) 
                  
              } # end of loop adding sites where full complement of sites were generates
           
           
            } else {
                 
                  print (paste0(i , " slice completed previously, skipping to next slice"))
                
                  prev_sample_points <- prev_sample_points %>%
                      dplyr::select(c(slice_num, geometry))
                  
                  # buffer the areas of this site for next sampling 
                    clhs_sampled_buff <- st_buffer(prev_sample_points, dist = rad_exclusion) 
                  
                      # Check if points overlap buffer and then remove 
                 # lays_df_SF <- lays_df  %>%
                #    st_as_sf(., coords = c("x","y")) %>%
                #    st_set_crs(3005)
                  
                  to_cost <- st_intersection(lays_sf, clhs_sampled_buff) %>%
                    dplyr::select(geometry) %>%
                    mutate(cost_update = 100000) 
                  
                  # this part is slow and could be sped up..
                  lays_df_buffered <- st_join(lays_sf, to_cost)  %>%
                    mutate(cost = ifelse(!is.na(cost_update), cost_update, cost)) %>%
                    dplyr::select(-cost_update) 
                
                  lays_sf <- lays_df_buffered  
                # lays_df <- cbind(st_coordinates(lays_df_buffered), lays_df_buffered) %>%
                #    st_drop_geometry() %>%
                #    dplyr::rename(x = X, y = Y)
          
                  sample_points <- bind_rows(sample_points, prev_sample_points)
                  
                  } # end of slice loop where no extra points needed. 


              } # end of slices loop  


  # convert to df to do full join
 sample_df <- sample_points_all_slice 
 sample_df <- cbind(sample_df, st_coordinates(sample_df)) %>%
   st_drop_geometry()
 
sample_points_df <- sample_points
sample_points_df <- cbind( sample_points_df , st_coordinates( sample_points_df )) %>%
   st_drop_geometry()
  
  xx <- full_join(sample_df, sample_points_df,  by = c("X", "Y")) %>%
    dplyr::select(-c(DAH, MRVBF, LFC, cost, slice_num.x)) %>%
    dplyr::rename("slice_num" = slice_num.y) %>%
    mutate(subzone = b, rotation = "cLHS") 

  sample_points_all <- xx %>%
  arrange(slice_num)

  

#sample_points
#   
# add the missing sites number for each slice
 out <- foreach(i = unique(sample_points_all$slice_num), .combine = rbind) %do%{
      #i = unique(sample_points_all$slice_num)[1]
      temp1 <- sample_points_all %>% filter(slice_num == i)
      missing.no = temp1$slice_size[!is.na(temp1$slice_size)]
      to.add <- dplyr::setdiff(seq(1:5), missing.no)
      temp1$slice_size[is.na(temp1$slice_size)] <- to.add
      temp1
  }

sample_points_all <- out %>%
#    sample_points_all <-  sample_points_all %>%
    arrange(slice_num, slice_size) %>%
   st_as_sf(coords = c("X","Y")) %>%
      st_set_crs(3005)

sample_points <- mutate(sample_points_all, total = 1:nrow(sample_points_all)) %>%
  mutate(id = paste0(subzone, "_", slice_num, ".", slice_size, "_", total)) %>%
  dplyr::select(names(.)[!(names(.) %in% names(lays))])

#rot = 3

 st_write(sample_points, dsn = file.path(clhs_outpath, paste0("clhs_pts_heli_",rot,"_", b,".gpkg")), driver = "GPKG")  
# #sample_points


#st_write(sample_points, dsn = file.path(clhs_outpath, "update_previous_sites", paste0("clhs_pts",rot,"_", b,".gpkg")), append = T, driver = "GPKG")  
#sample_points

  } # end loop of rotations

} # end of BGC loop 
    


```
