---
title: "QA for PEM field sampling Kootenay/Invermere/Cranbrook TSA 2022"
subtitle: "by Gen Perkins"
date: "22/11/2022"
output: html_document
---

```{r global_options, include=FALSE , warnings = FALSE, message = FALSE}
require(knitr)

```


```{r setup, include=FALSE, message = FALSE, warning = FALSE}

library(tidyverse)
library(raster)
library(fasterize)
library(sf)
library(tools)
library(lwgeom)
library(foreach)
library(tidyr)
library(purrr)
library(foreach)
library(scales)
library(stringr)
library(dplyr)
library(ggplot2)

```

```{r set up folder structure, include = FALSE, warning = FALSE }

#AOI <- "Deception"
AOI <- "KootInvCran"
res <- 5

AOI_dir <- file.path(".", paste0(AOI,"_AOI"))
cov_dir <- file.path(AOI_dir, "1_map_inputs", "covariates")

trans_layout <- file.path(AOI_dir, "2_sample_design", "stage1_StudyDesign", "transect_layout")
trans_data <- file.path(AOI_dir, "2_sample_design", "stage1_StudyDesign", "transect_data") 

output_pnts_dir <- file.path(AOI_dir, "2_sample_design", "stage1_StudyDesign", "training_pts")

qa_out_dir <- file.path(AOI_dir, "2_sample_design", "stage1_QA")
# trans_data <- file.path(AOI_dir, "2_sample_design","stage1_AA_2019")
# trans_layout <- file.path(AOI_dir, "2_sample_design","stage1_AA_2019")
# output <- file.path(AOI_dir, "3_maps_analysis", "external_AA")

map.key  <- read.csv(file.path(AOI_dir, "_MapUnitLegend", 
                                 paste0(AOI, "_MapUnitLegend.csv")), 
                       stringsAsFactor = FALSE)

res_folder <- paste0(res, "m")

raster_template  <- raster(file.path(cov_dir, "5m", "template.tif"))

if(!dir.exists(output_pnts_dir)) dir.create(output_pnts_dir, recursive = TRUE)
```

```{r Functions, include = FALSE}
source(here::here('_functions', 'extend_lines.R'))
source(here::here('_functions', 'make_lines.R'))
source(here::here('_functions', 'transect_sample.R'))
source(here::here('_functions', 'multiline_to_line.R'))
source(here::here('_functions', '_standardise_avenza_input_field_names.R'))
source(here::here('_functions', 'convert_line_allpts.R'))
#source(here::here('_functions', 'acc_metrix_QA.R'))

```

```{r import files to run analysis, tidy = TRUE,eval = FALSE, warning=FALSE, include = FALSE}

transect_layout <- st_read(file.path(output_pnts_dir, "transect_layout_s1.gpkg"))
transect_layout_buf <- st_buffer(transect_layout, 10)

```


```{r, include = FALSE, eval = FALSE}

observer_column <- "x02_observe"
poc_column <- "x03_pt_type"
mapunit1_column <- "x04_mapunit"
transition_column <- "x05_transit"
mapunit2_column <- "x06_mapunit"
strcst_column <- "x07_struct_"
strcmod_column <- "x08_struct_"
comment_column <- "x09_comment"
edatope_column <- "x10_edatope"
time_column <-"timestamp"

#order_column <- "f11_order"

points <- list.files(file.path(trans_data, "final_qa"), pattern = ".gpkg$|.shp$", full.names = TRUE)

#points <- points[1]


all_points <- foreach(x = points, .combine = rbind) %do% {

  #x = points[1]
  
  s1_layers <- st_layers(x)
  pts <- which(s1_layers[["geomtype"]] %in% c("Point","3D Point","3D Measured Point"))

  if(length(pts)>0) { 
    
  print(x)

     points_read <- st_read(x, quiet = TRUE) %>%
     #names(points_read)
     st_transform(3005) %>% 
     st_zm() %>% 
     rename_all(.funs = tolower) 
     
    if("x07_struc_" %in% names(points_read)){
        points_read <- points_read %>% 
          mutate(x07_struct_ = x07_struc_) %>%
          mutate(x08_struct_ = x08_struct)
    }
     
    if("f01_transec" %in% names(points_read)){
        dnames = names(points_read)
        colnames(points_read) <- gsub("f0", "x0", dnames)
        points_read <- points_read %>%
          dplyr::mutate(x10_edatope = f10_edatope)
          
      }
     

     points_read <- points_read %>% 
     mutate(mapunit1 = UQ(rlang::sym(mapunit1_column)),
            mapunit2 = UQ(rlang::sym(mapunit2_column)),
            point_type = UQ(rlang::sym(poc_column)),
            observer = UQ(rlang::sym(observer_column)), 
            transition = UQ(rlang::sym(transition_column)),
            comments = UQ(rlang::sym(comment_column)),
            strcst = UQ(rlang::sym(strcst_column)),
            strcmod = UQ(rlang::sym(strcmod_column)),
            time = UQ(rlang::sym(time_column))) %>%
     st_join(., transect_layout_buf, join = st_intersects) %>%
     rename_all(.funs = tolower) %>%
     rename(transect_id = id)
  
     if("name" %in% names(points_read)){
       points_read <- points_read %>%
         mutate(order = as.numeric(gsub("Placemark ", "", name)))
     }
     # } else {
     #   
     #   points_read <- points_read %>%
     #     mutate(name = seq(1, length(mapunit1),1))
     # }
    points_read <- points_read %>%
     dplyr::select(any_of(c("order", "mapunit1", "mapunit2", "point_type", "transect_id", "observer", "transition", "comments", "strcst", "strcmod", "time" ))) %>%
     group_by(transect_id) %>%
     arrange(as.numeric(order), by_group = TRUE) %>%
     ungroup() 
    
     st_geometry(points_read ) <- "geom"
     print(length(points_read$order))
     points_read
     
  }

} %>% 
  distinct(., .keep_all = TRUE) 

# catergorise data types:
points <- all_points %>%
  mutate(data_type = ifelse(is.na(transect_id), "incidental", "s1"))

# match to the baseunit key using mapunit legend & format observer column

# ## check the site names are matching in key 
# cross check values with AOI key
map.key_sum <- map.key %>% dplyr::select(c(FieldCall, BaseMapUnit))

# outdata <- points %>%
#   dplyr::select(mapunit1, mapunit2)%>%
#   st_drop_geometry() %>%
#   distinct() %>%
#   left_join(map.key_sum, by = c("mapunit1" = "FieldCall")) %>%
#   #dplyr::select(-mapunit1) %>%
#   dplyr::rename(mapunit1_match = BaseMapUnit) %>%
#   left_join(map.key_sum, by = c("mapunit2" = "FieldCall")) %>%
#   #dplyr::select(-mapunit2) %>%
#   dplyr::rename(mapunit2_match = BaseMapUnit) %>%
#   distinct()

points_final <- format_mapunit_names(points, map.key)
points_final <- fill_observer(points_final)
 
# covert to lines 
processed_transects <- make_lines(GPSPoints = points_final, 
                                  Transects = transect_layout_buf, 
                                  method = "pts2lines",  
                                  tBuffer = 20, PROJ = 3005) %>%
  mutate(mapunit1 = (trimws(mapunit1)),
         mapunit2 = (trimws(mapunit2))) %>%
  mutate(mapunit12 = paste0(mapunit1,"_", mapunit2)) %>%
  mutate(mapunit12 = gsub("_NA","", mapunit12)) %>%
  mutate(mapunit12 = gsub("_$","", mapunit12)) %>%
  dplyr::select(-TID, -ID)

st_write(processed_transects,  file.path(output_pnts_dir, "QA_proc_s1_transects.gpkg"), 
         delete_layer = TRUE)

st_write(points_final, file.path(output_pnts_dir, "QA_s1_pts.gpkg"), delete_layer = TRUE)

```

```{r import base files to run analysis, tidy = TRUE, warning=FALSE, eval = FALSE, include = FALSE}
## 1) Import consolidated data and linear transects

transect_layout <- st_read(file.path(output_pnts_dir, "transect_layout_s1.gpkg"))
transect_layout_buf <- st_buffer(transect_layout, 10)


s1_trans_QA <- st_read(file.path(output_pnts_dir, "QA_proc_s1_transects.gpkg"), quiet = TRUE) %>% mutate(data_type ="QA")
raster_points_xy <- convert_line_allpts(s1_trans_QA, terra::rast(raster_template))
raster_points_xy <- format_transect(st_as_sf(raster_points_xy))
st_write(raster_points_xy, file.path(output_pnts_dir, "QA_transect_all_pts.gpkg"), 
                              delete_layer = TRUE)
      

# filter the contractor data that has been QA's
s1_trans_cont  <- st_read(file.path(output_pnts_dir, "s1_transects.gpkg")) %>% mutate(data_type ="contractor") 
raster_points_xy <- convert_line_allpts(s1_trans_cont , terra::rast(raster_template))
raster_points_xy <- format_transect(st_as_sf(raster_points_xy))
st_write(raster_points_xy, file.path(output_pnts_dir, "cont_transect_all_pts.gpkg"), 
         delete_layer = TRUE)

```
 
 
```{r, echo = FALSE, fig.width = 10, fig.height=10, message = FALSE}

## Summary 
qp <- st_read(file.path(output_pnts_dir, "QA_transect_all_pts.gpkg"), quiet = T) %>%
   cbind(st_coordinates(.)) %>% st_drop_geometry()

cp <- st_read(file.path(output_pnts_dir, "cont_transect_all_pts.gpkg"), quiet = T) %>% cbind(st_coordinates(.)) %>% st_drop_geometry()

# combine the QA and cp points
allpts <- left_join(qp, cp, by = c("X", "Y"))

# set up table with mapunit1_QA and mapunit1_co (only for overlapping ground)
subpts <- allpts %>%
  dplyr::select(transect_id.x, tid.x, mapunit1.x, mapunit2.x, transition.x, observer.x, comments.x , mapunit1.y,  mapunit2.y, transition.y, observer.y,comments.y , X, Y) %>%
  filter(!is.na(mapunit1.y)) 


sb <- st_as_sf(subpts, coords = c("X","Y"), crs = st_crs(3005))


          
```



```{r, eval = TRUE, include = FALSE}

# number of transects 
ssb <- subpts %>% 
  filter(!is.na(transect_id.x)) %>% 
  filter(!is.na(mapunit1.x)) %>%
  filter(!is.na(observer.y)) %>%
  filter(transect_id.x != "MS  dw_2.5_10_cLHS") %>%
  filter(transect_id.x != "IDF dm 2_2.5_10_cLHS")

unique(ssb$transect_id.x)
unique(ssb$observer.y)

# update the observer name to make annonymous

ssb <- ssb %>% 
  mutate(observer.y = case_when(
    observer.y == "BB" ~ "junior_1",
    observer.y == "DK" ~ "intermediate_1",
    observer.y == "MvW" ~ "senior_1",
    observer.y == "SH" ~ "senior_2",
  ))



# 18 transects 
```

In 2022, quality assurance surveys were completed on `r length(unique(ssb$transect_id.x))` transects from `r length(unique(ssb$observer.y))` observers. 

Only locations (pixals at 5m scale) which were directly overlapping were used to compare between the QA and contractors. The total number of overlapping pixals was `r length(ssb$observer.y)`

```{r, include = FALSE, eval = TRUE}
# number of observers 
obs_sum <- ssb %>%
  filter(!is.na(observer.y)) %>%
  group_by(observer.y, transect_id.x) %>%
  summarise(no = n())

obs <- ssb %>%
  filter(!is.na(observer.y)) %>%
  group_by(observer.y) %>%
  summarise(no = n()) 

trans <- ssb %>%
  filter(!is.na(observer.y)) %>%
  group_by( transect_id.x) %>%
  summarise(no = n()) 

```


```{r, eval = TRUE}
knitr::kable(obs_sum)
```


This included a mix of subzone variants. 
I compared both primary and secondary call for both contractor and QA calls to give benefit of doubt for calls. 


```{r, eval = TRUE, include = FALSE}

qa_summary <- ssb %>% 
  dplyr::select(transect_id.x, mapunit1.x,  mapunit2.x,
                observer.y,mapunit1.y, mapunit2.y)

qa1 <- qa_summary %>%
  mutate(best_ss = case_when( 
    mapunit1.y == mapunit1.x ~ mapunit1.y, 
    mapunit1.y != mapunit1.x & mapunit2.y == mapunit1.x ~ mapunit2.y,
    mapunit1.y != mapunit1.x & mapunit1.y == mapunit2.x ~ mapunit1.y
    )) %>%
  mutate(final_ss = case_when(
    is.na(best_ss) ~ mapunit1.y, 
    TRUE ~ best_ss))
    
qamatrix <- dplyr::select(qa1, transect_id.x, observer.y, mapunit1.x, final_ss)


# format levels for factors 
levels = with(qamatrix, unique(c(mapunit1.x, final_ss)))

qamatrix = within(qamatrix, {
    mapunit1.x = factor(mapunit1.x, levels)
  })
qamatrix = within(qamatrix, {
    final_ss = factor(final_ss, levels)
  })

names(qamatrix) = c("transect", "observer", "target", ".pred_class")

uids <- as.factor(unique(qamatrix$transect))
  
# for all slices
sresults <- foreach(k = levels(uids)) %do% {
  
  #k = levels(uids)[1]
  print(k)

  # test set
  tdata <- qamatrix %>% filter(transect %in% k)
  
  # calculate the aspatial metrics: 
  
  aspatial_pred <- tdata  %>% 
    dplyr::select(.pred_class) %>% 
    group_by(.pred_class) %>% 
    dplyr::mutate(pred.tot = n()) %>% 
    ungroup() %>% distinct()
  
  aspatial_target <- tdata %>% 
    dplyr::select(target) %>% 
    group_by(target) %>% 
    dplyr::mutate(trans.tot = n()) %>% 
    ungroup() %>% distinct()
  
  aspatial_sum <- full_join(aspatial_target, aspatial_pred, by = c("target" = ".pred_class")) %>% 
    mutate_if(is.integer, funs(replace_na(., 0))) %>% 
    mutate(trans.sum = sum(trans.tot, na.rm = TRUE)) %>% 
    dplyr::rowwise() %>% 
    mutate(trans.pro = (trans.tot/trans.sum), 
           pred.pro = (pred.tot/trans.sum)) %>%
    ungroup() 
  
  # work around to fix min() rowwise error
    dat <- data.table::as.data.table(aspatial_sum)
    dat[, aspat_p := pmin(trans.pro, pred.pro)]
    dat[, unit_pos_temp := pmin(trans.tot, pred.tot)]
    
    #mutate(unit_pos = min(trans.tot, pred.tot)/trans.tot) %>%
    
  aspatial_sum = as.data.frame(dat) %>%
    dplyr::select(c(-trans.pro, -pred.pro)) %>%
    mutate(aspat_p_mean = sum(aspat_p)) %>%
    dplyr::rowwise() %>%
    mutate(unit_pos = unit_pos_temp/trans.tot)%>%
    dplyr::select(-unit_pos_temp) %>%
    mutate(transect = unique(tdata$transect),
           observer = unique(tdata$observer))
    aspatial_sum[is.na(aspatial_sum)] <- 0

  
  # calculate the spatial metrics
  
  test.pred <- dplyr::select(tdata, "target", ".pred_class")
  
  ###harmonize levels
  targ.lev <- levels(test.pred$target)
  pred.lev <- levels(test.pred$.pred_class)
  levs <- c(targ.lev, pred.lev) %>% unique()
  test.pred$target <- factor(test.pred$target, levels = levs)
  test.pred$.pred_class <- factor(test.pred$.pred_class, levels = levs)

  cfm <- as_tibble(table(test.pred)) 
  cfm <- cfm %>% 
    filter(n >0) %>% 
    mutate(transect = unique(tdata$transect),
           observer = unique(tdata$observer))
  
  spatial <- cfm %>% 
    dplyr::mutate(keep = case_when(
      target == .pred_class ~ 1
    )) %>% 
    filter(keep == 1) %>% 
    mutate(spat_no = as.numeric(n)) %>% 
    dplyr::select(target, spat_no)

  tot_acc <- dplyr::left_join(aspatial_sum, spatial, by = "target")
  tot_acc [is.na(tot_acc )] <- 0
  
  tot_acc <- tot_acc %>%
    rowwise() %>%
    mutate(spat_p_unit = round(spat_no/trans.tot,3)) %>%
    ungroup()
  tot_acc [is.na(tot_acc )] <- 0
  
   tot_acc <- tot_acc %>%
     mutate(spat_sum = sum(spat_no)) %>% 
     mutate(spat_mean = round(spat_sum/trans.sum,2)) %>% 
     dplyr::select (-aspat_p)
  
  return(list(cfm,  tot_acc ))

  }


# extract results from sresults
cov_matrix <- lapply(sresults, function(x) x[[1]])
tots <- lapply(sresults, function(x) x[[2]])
cov_matrix <- as.data.frame(data.table::rbindlist(cov_matrix)) 
tots <- as.data.frame(data.table::rbindlist(tots)) 
  
names(cov_matrix)<- c("qa", "obs", "n", "transect", "observer")

```

    
## Aggreement measures

Several types of agreement measures were calculated;

1) unweighted (aspatial): this is equivalent to traditional AA where proportion of map units are compared. 

2) area weighted (spatial) (spat_p): this compares spatial equivalents for best call (primary and or secondary) for each pixal/point predicted.



## Aspatial and Spatial Agreement

```{r, echo = FALSE, eval = TRUE}
#Compare the CV metrics to test metrics to see model fit 

acc_sum <- tots  %>%
    dplyr::mutate(across(ends_with("mean"), ~.x *100)) %>%
    #dplyr::mutate(across(ends_with("meanacc"), ~.x *100)) %>%
    dplyr::select(transect, observer, target,
                  aspat_p_mean, spat_mean,
                  unit_pos, spat_p_unit)

average_acc <- acc_sum %>%
   dplyr::select(transect, observer,
                  aspat_p_mean, spat_mean) %>%
  distinct()
  

mapunit_acc <- acc_sum %>%
   dplyr::select(transect, observer, target,
                  unit_pos, spat_p_unit) %>%
  distinct()
  
  
# plot both the cv and test metrics
p1 <- ggplot(aes(y = aspat_p_mean, x = transect ), data = average_acc ) + 
   geom_boxplot() +
   facet_wrap(~observer, scales = "free_x", nrow = 1) +
   geom_hline(yintercept = 65,linetype ="dashed", color = "black") + 
   ggtitle("Mean Aspatial Agreement") + 
   xlab("Mapunit") + ylab("Accuracy") + 
   ylim(-0.05, 100) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
   #scale_fill_discrete_sequential(palette = "Light Grays")

p1


# plot both the cv and test metrics
p2 <- ggplot(aes(y = spat_mean, x = transect ), data = average_acc ) + 
   geom_boxplot() +
   facet_wrap(~observer, scales = "free_x", nrow = 1) +
   geom_hline(yintercept = 65,linetype ="dashed", color = "black") + 
   ggtitle("Mean Spatial Agreement") + 
   xlab("Mapunit") + ylab("Accuracy") + 
   ylim(-0.05, 100) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
   #scale_fill_discrete_sequential(palette = "Light Grays")

p2

```


## Agreement per mapunit

We can compare map unit agreement levels to assess under or acceptable performance per map units. 

```{r generate overall mapunit, echo = FALSE, eval = TRUE}

# map unit plots: 

mu_acc <- mapunit_acc %>%
  pivot_longer(cols = where(is.numeric), names_to = "accuracy_type", values_to = "value") 

mu_acc <- mu_acc %>%
  mutate(type = case_when(
    str_detect(accuracy_type, "unit_pos") ~ "aspatial",
    str_detect(accuracy_type, "spat") ~ "spatial"))

p4 <- ggplot(aes(y = value, x = target , colour = type), data = mu_acc  ) + 
   geom_jitter() +
   facet_wrap(~observer, scales = "free_x", nrow = 1) +
   ggtitle("Mapunit accuracy measures ") + 
   xlab("accuracy measure") + ylab("Proportion of Accurate calls") + 
   ylim(-0.05, 1) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
p4

p5 <- ggplot(aes(y = target, x = value , colour = observer), data = mu_acc  ) + 
   geom_boxplot() +
   facet_wrap(~type, scales = "free_x", nrow = 1) +
   ggtitle("Mapunit accuracy measures ") + 
   ylab("Mapunit") + xlab("Proportion of Accurate calls") + 
   #ylim(-0.05, 1) +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
p5


```



## Confusion matrix

```{r, include = TRUE}
#knitr::kable(cov_matrix)
# 
# # testing for across matrix
uids <- as.factor(unique(cov_matrix$transect))

for(i in uids){
 # i <- uids[14]
  
  print(i)
  
  cm <- cov_matrix %>% 
     dplyr::filter(transect == i) %>%
    dplyr::select(-transect) %>% 
    dplyr::filter(qa !="")

  xx <-  cm %>%
   pivot_wider(names_from = qa, values_from = n, values_fill = 0,values_fn = ~sum(.x, na.rm = TRUE)) %>% dplyr::relocate(observer)

  print(xx)
  #knitr::kable(xx)
} 

# xx <-  cm %>%
#   pivot_wider(names_from = qa, values_from = n, values_fill = 0,values_fn = ~sum(.x, na.rm = TRUE))
knitr::kable(cov_matrix)
```



```{r, echo = FALSE, fig.width= 12, fig.height = 16, eval = FALSE}
# plot the freqency 

# points collected per type (primary and secondary calls) all calls
#p <- ggplot(trans_total, aes(y = mapunit12, x = length)) + 
#  facet_wrap(~bgc, scales = "free_x") +    
#  geom_bar(stat = "identity")
#         
#print(p)



for(b in bgc_list) {

 # b = unique(pt_total$bgc)[3]
  pt_total_bgc <- trans_total %>%
    filter(bgc == b)
  
  p <- ggplot(pt_total_bgc , aes(y = reorder(mapunit12, length), x = length)) +
  geom_bar(stat = "identity") + 
  labs(title = b, y = "mapunit")
  
  print(p)
  
#  ggplot(pt_total_bgc , aes(y = mapunit1, x = freq)) +
#  geom_bar(stat = "identity")
     
  #ggsave(file.path(s1_dir, paste0(b, "_point_freq_s1.png")))
  
  
}

```


```{r, echo = FALSE, eval = FALSE}

# group by observer

all_trans_ob <- all_trans %>%
  group_by(transect_id) %>%
  fill( observer, .direction = "downup")

#all_trans_ob_sum <- all_trans_ob %>%
#   group_by(transect_id, observer) %>%
#    dplyr::summarise(count = n())

trans_ob_p <- all_trans_ob %>%
 # dplyr::filter(is.na(mapunit2)) %>%
  group_by(observer) %>%
  summarise(all.calls = sum(length))

trans_ob_s <- all_trans_ob %>%
  dplyr::filter(!is.na(mapunit2)) %>%
  group_by(observer) %>%
  summarise(sec.calls = sum(length))


trans_obs_type <- left_join(trans_ob_p, trans_ob_s) %>%
  mutate(pc.sec = round((sec.calls/all.calls * 100),1))

trans_obs_type<- trans_obs_type %>%
  mutate(observer = case_when(
    observer == "EAC GP" ~ "EAC",
    observer == "EAC GCP" ~ "EAC",
    observer == "EC" ~ "EAC",
    observer == "CC LJ" ~ "CC",
    observer == "DF LJ" ~ "DF",
    observer =="DJM DF" ~ "DJM",
    observer == "PD AR" ~ "PD", 
    observer =="SRan" ~ "SR",
    observer == "SR/MZ" ~ "SR",
    observer == "WM" ~ "WHM",
        TRUE ~ observer)) %>%
      group_by(observer) %>%
      summarise(across(where(is.numeric), sum)) %>%
  mutate(pc.sec = round(sec.calls/all.calls * 100,1))

```

```{r, results  = "asis", hide = TRUE, echo = FALSE, eval = FALSE}

kable(trans_obs_type)

```


