---
title: "PEM_ColourThemeing"
author: "Kara Schaefer, Gen Perkins"
date: "11/08/2019"

output:
  pdf_document: default
  html_document: default
  word_document: default
---
<!--
Copyright 2020 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(readr)
library(ggplot2)
library(XML)
library(colortools)
library(sf)
library(lwgeom)
library(sp)
library(dplyr)
library(bcdata)
library(bcmaps)
library(readr)


```
##Script to assign colours to map units in a PEM based on complementary colours.
The user will read in a Map Unit Legend and this script will assign and write colours back to the legend files.  
The zonal site series colour will match the colours used for the BGC subzone variant
Then complementary, adjacent, or analagous colours will be selected based on edatopic position
Special colours/patterns will be assigned to non-forested classes.
Some links to script descriptions: 
https://www.datanovia.com/en/blog/how-to-create-beautiful-color-palettes-in-r/

Online colour tool
http://colorschemedesigner.com/csd-3.5/


```{r import BGC units and map_keys, tidy = TRUE, warning=FALSE}

AOI <- "Deception"
AOI_dir <- file.path(paste0(AOI,"_AOI"))
map_unit_legend <- file.path(AOI_dir, "_MapUnitLegend")
out_dir <- file.path(AOI_dir, "3_maps_analysis","models", "forest","forest_mu_all", "77")

qml_out <- file.path(out_dir, "map")

# read in reference files 
map_key  <- read.csv(file.path(AOI_dir, "_MapUnitLegend", 
                                 paste0(AOI, "_MapUnitLegend.csv")), 
                       stringsAsFactor = FALSE)


# read in raster file and map key 
map_title = "response_names"

indata <- read_csv(file.path(out_dir, paste0(map_title, ".csv"))) %>%
  rename(category = x,
         integercode = X1)

# prepare the edatopic grid reference if needed 

if(file.exists(file.path(map_unit_legend, "edatopic_reference.csv"))) {
  print("reading existing edatopic_reference")
  
  edatopic_reference = read.csv(file.path(map_unit_legend, 
                                          "edatopic_reference.csv"), 
                                stringsAsFactors = FALSE)
  SUBZONE_COLOURS_KEY = read.csv(file.path(map_unit_legend, 
                                           "subzone_colour_key.csv"), 
                                 stringsAsFactors = FALSE)
  
} else {
  print("generating edatopic reference")
  
  if(file.exists(file.path(map_unit_legend, "Edatopic_v11_19.csv"))) {
    
    
    
    edatopic_reference <- read.csv(file.path(map_unit_legend, "Edatopic_v11_19.csv"))
    
    file_name <- "WNA_SubzoneColours"
    
    document <- xmlParse(file = file.path(map_unit_legend, paste(file_name, ".xml", sep = "")))
    category_nodes <- getNodeSet(document, "//category")
    BGC_zone <- lapply(category_nodes, xmlGetAttr, "value")
    xml_id <- lapply(category_nodes, xmlGetAttr, "symbol")
    BGC_with_xml_id_only <- data.frame(BGC_zone = unlist(BGC_zone), xml_id = unlist(xml_id))
    
    symbol_nodes <- getNodeSet(document, "//symbols//symbol[@name]")
    symbol_id <- sapply(symbol_nodes, xmlGetAttr, "name")
    names(symbol_id) <- sapply(getNodeSet(document, '//symbols//symbol//prop[@k="color"]'), xmlGetAttr, "v")
    
    id_and_colour <- data.frame(id = symbol_id,
                                cmyk_colour = names(symbol_id), stringsAsFactors = FALSE)
    
    BGCs_with_colours <- BGC_with_xml_id_only %>% 
      left_join(id_and_colour, by = c("xml_id" = "id"))
    
    BGCs_with_colours <- BGCs_with_colours[, c("BGC_zone","cmyk_colour", "xml_id") ]
    
    
    SUBZONE_COLOURS_KEY <- BGCs_with_colours %>% 
      mutate(colour = sapply(
        strsplit(BGCs_with_colours$cmyk_colour, ","),
        function(x) rgb(x[1], x[2], x[3], maxColorValue = 255)),
        classification = paste(BGC_zone, "/01", sep = "")) %>%
      dplyr::select(classification, colour) %>%
      rbind(c("ESSFmcw/101", "#6B0387"))    # manually add a purple for ESSFmcw, as it is missing
    
    
    # manually add ESSFmcw edatopic information
    ESSFmcw_edatopic <- data.frame(
      "--",
      "ESSFmcw",
      c(rep("ESSFmcw/101", 4),
        rep("ESSFmcw/102", 4),
        rep("ESSFmcw/103", 5),
        rep("ESSFmcw/110", 8),
        rep("ESSFmcw/111", 4)),
      c(c("B3", "B4", "C3", "C4"),
        c("A1", "B1", "B2", "C2"),
        c("A2", "A3", "A4", "B3", "B4"),
        c("B5", "C5", "D3", "D4", "D5", "E3", "E4", "E5"),
        c("B6", "C6", "D6", "E6")),
      NA,
      NA,
      NA,
      NA,
      NA,
      NA,
      NA,
      NA
    )
    
    names(ESSFmcw_edatopic) <- colnames(edatopic_reference)
    edatopic_reference <- rbind(edatopic_reference, ESSFmcw_edatopic)
    write.csv(edatopic_reference, file.path(map_unit_legend, "edatopic_reference.csv"), row.names = 
                FALSE)
    write.csv(SUBZONE_COLOURS_KEY, file.path(map_unit_legend, "subzone_colour_key.csv"), row.names = FALSE)
  } else {
    print("Edatopic_v11_19.csv file not found in map_unit_legend directory")
  }
}


#BGCs <- read.csv(file.path(map_unit_legend, "AllModelled_BGCs_v11.csv")) ### BGC RGB colour will be assigned here manually picked to match the BGC field maps
#BGCs <- BGCs[!is.na(BGCs),]


# alternative to use bc catalogue or another dataset with unique bgc 
#bec_sf <- bec(class = "sf")
#bec_sf <- bcdc_query_geodata("0d526b63-812b-444c-b227-418b6fcbbd44")%>%
#      bcdata::select(c("BG_CODE")) %>% # Treed sites
#      collect()

#examples <- file.path(map_unit_legend, "example_map_keys")
#map_title <- "/M57_map_key.csv"
#indata <- read.csv(file.path(examples, map_title), stringsAsFactors = FALSE)

# indata <- read.csv(file.path(examples, "/M28_map_key.csv"), stringsAsFactors = FALSE)
# indata <- read.csv(file.path(examples, "/M37_map_key.csv"), stringsAsFactors = FALSE)
#M42_example <- read.csv(file.path(examples, "/M42_map_key.csv"), stringsAsFactors = FALSE)
#M55_example <- read.csv(file.path(examples, "/M55_map_key.csv"), stringsAsFactors = FALSE)


# Set up colour tables for set non-forest and forest groups 

# TO DO: 
# Non-forest perhaps set up a pattern (not just a solid colour, hashes or lines) - colour by theme ie shades of blue for common classes 
# ie Wf01 Wf02
# Forest - contrasting (not theme shaded) - set up the manual 
# see the arcmap set themeing for specific - ie wetlands. 

nf_colour_table <- tribble( #nf = non-forest
  ~classification, ~colour,
  "LA",       "#0000cc",    # Lakes, bold blue
  "RI",       "#0000cc",    # Rivers, bold blue
  "Unknown",  "#808080"     # grey
)

id <- c(
  paste0(toupper(letters[1:5]), 0),
  paste0(toupper(letters[1:5]), 1),
  paste0(toupper(letters[1:5]), 2),
  paste0(toupper(letters[1:5]), 3),
  paste0(toupper(letters[1:5]), 4),
  paste0(toupper(letters[1:5]), 5),
  paste0(toupper(letters[1:5]), 6),
  paste0(toupper(letters[1:5]), 7)
)

position_value <- 1:40

edatopic_position_values <- data.frame(id, position_value, stringsAsFactors = FALSE)

imported_table_full <- indata %>% 
  mutate(
    category = ifelse(grepl("_", category), 
                      sub("_", "/", category, perl = TRUE),
                      category),
    forest_non_forest = ifelse(
      grepl("(?i)/[a-z]", category, perl = TRUE) | !grepl("/", category, perl = TRUE),
      "non_forest",
      "forest"),
    category_mapped_for_colour = ifelse(
      grepl("(?i)/[a-z]", category, perl = TRUE),
      sub(".*/", "", category),
      category),
    category_mapped_for_colour = ifelse(
      category_mapped_for_colour %in% c("Un", "Unk", "un", "?"),
      "Unknown",
      category_mapped_for_colour
    ),
    # Remove numeric values from non-forested types
    # for example Wf01 becomes Wf
    category_mapped_for_colour = ifelse(
      grepl("non_forest", forest_non_forest, perl = TRUE),
      ifelse(
        grepl("[0-9]", category_mapped_for_colour, perl = TRUE),
        sub("^[0-9 ,]+|[0-9 ,]+$", "", category_mapped_for_colour),
        category_mapped_for_colour
      ),
      category_mapped_for_colour
    ),
    mapped_to = case_when(
      grepl("(?i)unknown", category_mapped_for_colour, perl = TRUE) ~ "Unknown",
      grepl("(?i)ag", category_mapped_for_colour, perl = TRUE) ~ "grassland",
      grepl("(?i)am", category_mapped_for_colour, perl = TRUE) ~ "meadow",
      grepl("(?i)es$", category_mapped_for_colour, perl = TRUE) ~ "estuarine_maybe",
      grepl("(?i)fl", category_mapped_for_colour, perl = TRUE) ~ "flood_group",
      grepl("(?i)ro", category_mapped_for_colour, perl = TRUE) ~ "rock_outcrop",
      grepl("(?i)sc", category_mapped_for_colour, perl = TRUE) ~ "shrubland",
      grepl("(?i)wa", category_mapped_for_colour, perl = TRUE) ~ "alpine_wetland",
      grepl("(?i)wb", category_mapped_for_colour, perl = TRUE) ~ "bog",
      grepl("(?i)wf", category_mapped_for_colour, perl = TRUE) ~ "fen",
      grepl("(?i)wm", category_mapped_for_colour, perl = TRUE) ~ "marsh",
      grepl("(?i)wun", category_mapped_for_colour, perl = TRUE) ~ "wu_n",
      grepl("(?i)ww", category_mapped_for_colour, perl = TRUE) ~ "shallow_water",
      grepl("(?i)xvh", category_mapped_for_colour, perl = TRUE) ~ "disclimax_avalanche_herb_meadow",
      grepl("(?i)xvs", category_mapped_for_colour, perl = TRUE) ~ "disclimax_avalanche_shrub_thicket",
      grepl("(?i)xvt", category_mapped_for_colour, perl = TRUE) ~ "disclimax_avalanche_treed"
    ),
    variant = ifelse(
      forest_non_forest == "forest",
      sub("/.*", "", category),
      NA
    )
  )

forest_present <- ifelse(
  nrow(imported_table_full %>% 
         filter(forest_non_forest == "forest")) > 0,
  TRUE,
  FALSE)

non_forest_present <- ifelse(
  nrow(imported_table_full %>% 
         filter(forest_non_forest == "non_forest")) > 0,
  TRUE,
  FALSE)

# custom method to check that a value is _not_ present in a vector
"%!in%" <- function(x, table) match(x, table, nomatch = 0) == 0

warning_list <- NA

remove_FF <- function(colour_param) {
  # in certain variations of the colortools::sequential method, it adds a trailing FF on top of the standard 6 character hexadecimal colour. This function checks for FF values at the end of the string and then removes any characters after the regular hexadecimal code
  
  x <- as.list(colour_param)
  if(grepl(".*FF$", x, perl = TRUE)){
    colour_param <- unlist(
      lapply(x, function(i){
        j <- substr(i, 1, 7)
      })
    )
  }
  return(colour_param)
}

if(non_forest_present) {
  starting_colour <- "#808080"
  non_forest_table <- imported_table_full %>% 
    filter(
      forest_non_forest == "non_forest",
      category_mapped_for_colour %!in% nf_colour_table$classification
    ) %>% 
    select(category_mapped_for_colour) %>% 
    distinct() %>% 
    mutate(colour_categories = nrow(.),
           percent = case_when(
             colour_categories == 1 ~ 0,
             colour_categories == 2 ~ 40,
             colour_categories == 3 ~ 33,
             colour_categories == 4 ~ 25,
             colour_categories == 5 ~ 20,
             colour_categories == 6 ~ 15,
             colour_categories == 7 ~ 13,
             colour_categories == 8 ~ 12,
             colour_categories == 9 ~ 11,
             colour_categories == 10 ~ 10,
             colour_categories == 11 ~ 9,
             colour_categories == 12 ~ 8,
             colour_categories == 13 ~ 7,
             colour_categories == 14 ~ 7,
             colour_categories == 15 ~ 6,
             colour_categories == 16 ~ 6,
             colour_categories == 17 ~ 5,
             colour_categories == 18 ~ 5,
             colour_categories > 19 ~ 2,
           ))
  
  percent_non_forested <- unique(non_forest_table$percent)
  non_forested_colours <- sequential(starting_colour,
                                     what = "value",
                                     percentage = percent_non_forested,
                                     fun = "sqrt",
                                     plot = TRUE)
  
  # the first colour is always black so we shall remove it
  # however we need to calculate the length of the colour list specially because of the sequential percentage when there are potentially a large number of nf_colours
  colour <- non_forested_colours[2:(nrow(non_forest_table) + 1)]
  colour <- remove_FF(colour)
  
  classification <- non_forest_table %>% 
    select(category_mapped_for_colour) %>% 
    pull
  
  nf_colour_table <- rbind(
    nf_colour_table,
    data.frame(
      classification, colour, stringsAsFactors = FALSE)
  )
}

if(forest_present) {
  forested_BGCs <- imported_table_full %>% 
    filter(forest_non_forest == "forest") %>% 
    select(category) %>% 
    distinct() %>% 
    pull
  
  unique_BGCs <- unique(sub("/.*", "", forested_BGCs))
  
  zonal_colours <- SUBZONE_COLOURS_KEY %>% 
    filter(
      classification %in% imported_table_full$category
    ) %>% 
    mutate(
      BGC = sub("/.*", "", classification)
    )
  
  for(i in unique_BGCs) {
    # i = unique_BGCs[1]
    
    zonal_colour <- zonal_colours %>% 
      filter(
        BGC == i
      ) %>% 
      select(colour) %>%
      pull()
    
    if(length(zonal_colour) < 1) {
      warn_01 <- warning(paste0('zonal colour not found for ', i, ', assigning #FF6347'))
      warning_list <- as.list("zonal_colour": warn_01)
      print(warn_01)
      zonal_colour <- "#FF6347"
    }
    
    imported_table <- imported_table_full %>% 
      filter(variant == i,
             forest_non_forest == "forest")
    
    imported_table <- imported_table %>% 
      left_join(edatopic_reference, by = c("category" = "SS_NoSpace")) %>% 
      mutate(
        site_series = ifelse(
          grepl("(?i)/[0-9]", category, perl = TRUE),
          sub(".*/", "", category),
          category),
        x = case_when(
          grepl("(?i)A", Edatopic, perl = TRUE) ~ "A",
          grepl("(?i)B", Edatopic, perl = TRUE) ~ "B",
          grepl("(?i)C", Edatopic, perl = TRUE) ~ "C",
          grepl("(?i)D", Edatopic, perl = TRUE) ~ "D",
          grepl("(?i)E", Edatopic, perl = TRUE) ~ "E",
        ),
        y = case_when(
          grepl("0", Edatopic) ~ "0",
          grepl("1", Edatopic) ~ "1",
          grepl("2", Edatopic) ~ "2",
          grepl("3", Edatopic) ~ "3",
          grepl("4", Edatopic) ~ "4",
          grepl("5", Edatopic) ~ "5",
          grepl("6", Edatopic) ~ "6",
          grepl("7", Edatopic) ~ "7",
        )
      )
    
    imported_table <- imported_table %>% 
      left_join(edatopic_position_values, by = c("Edatopic" = "id"))
    
    x.ls <- as.list(forested_BGCs)
    out <- lapply(x.ls, function(xy) {
      temp <- imported_table %>% 
        filter(category == xy) %>% 
        select(position_value) %>% 
        filter(position_value == min(position_value)) %>% 
        distinct() %>% 
        mutate(category = xy,
               minimum_position_value = position_value) %>% 
        select(category, minimum_position_value)
    })
    out <- do.call("rbind", out)
    
    imported_table <- imported_table %>%
      left_join(out)
    imported_table <- imported_table[, c("category", "Edatopic", "site_series", "x", "y", "position_value", "minimum_position_value", "forest_non_forest", "variant", "category_mapped_for_colour", "mapped_to")]
    
    edatopic_position_01_minimum <-
      imported_table %>% 
      filter(site_series %in% c("01", "101")) %>% 
      select(minimum_position_value) %>% 
      distinct() %>% 
      pull()
    
    categories_with_cooler_colours <- imported_table %>% 
      filter(as.numeric(minimum_position_value) < edatopic_position_01_minimum) %>% 
      select(category, minimum_position_value) %>% 
      distinct() %>% 
      mutate(colour_categories = nrow(.),
             percent = case_when(
               colour_categories == 1 ~ 0,
               colour_categories == 2 ~ 40, # the higher the percent, the fewer colours will be in the colour palette. Fewer are needed with fewer colour categories.
               colour_categories == 3 ~ 33,
               colour_categories == 4 ~ 25,
               colour_categories == 5 ~ 20,
               colour_categories == 6 ~ 15,
             ))
    
    categories_with_warmer_colours <- imported_table %>% 
      filter(as.numeric(minimum_position_value) > edatopic_position_01_minimum) %>% 
      select(category, minimum_position_value) %>% 
      distinct() %>% 
      mutate(colour_categories = nrow(.),
             percent = case_when(
               colour_categories == 1 ~ 0,
               colour_categories == 2 ~ 40,
               colour_categories == 3 ~ 33,
               colour_categories == 4 ~ 25,
               colour_categories == 5 ~ 20,
               colour_categories == 6 ~ 15,
             ))
    
    if(length(categories_with_cooler_colours) && length(categories_with_warmer_colours)) {
      three_starting_points <- triadic(zonal_colour)
    }
    
    x <- col2rgb(three_starting_points[2])
    y <- col2rgb(three_starting_points[3])
    # The hottest colour is completely red (R=255, G=0, B=0).
    # The coldest colour is completely blue (R=0, G=0, B=255).
    # So the more R you have, the warmer the colour, and the more B you have, the cooler the colour.
    # after some discussion with the team, it was determined to flip this logic about warm/cold.
    if(x[1] < y[1]) {
      warm_starting_colour <- three_starting_points[2]
      cool_starting_colour <- three_starting_points[3]
    } else {
      warm_starting_colour <- three_starting_points[3]
      cool_starting_colour <- three_starting_points[2]
    }
    
    percent_cool <- unique(categories_with_cooler_colours$percent)
    cool_colours <- sequential(cool_starting_colour,
                               what = "value",
                               percentage = percent_cool,
                               v = 1,
                               alpha = 1,
                               fun = "sqrt",
                               plot = TRUE)
    cool_colours <- rev(cool_colours[-1])
    cool_colours <- remove_FF(cool_colours)
    
    categories_with_cooler_colours <- categories_with_cooler_colours %>% 
      mutate(colour = cool_colours)
    
    # if there are more than three colours needed for a colour palette, start a whole new one so that the overall colour theme for dryer/warmer colours is more varied. For example in the dryer colour scheme for a forest type, you might have three yellow shades and two purple shades, but all five represent dryer colours.
    if(nrow(categories_with_warmer_colours) > 3) {
      first_warm_colour <- colortools::adjacent(warm_starting_colour)[3]
      second_warm_colour <- colortools::adjacent(warm_starting_colour)[2]
      percent_warm <- 33
      warm_colours_1 <- sequential(first_warm_colour,
                                   what = "value",
                                   percentage = percent_warm,
                                   plot = TRUE)
      
      warm_colours_1 <- rev(warm_colours_1[2:length(warm_colours_1)])
      colours_left <- nrow(categories_with_warmer_colours) - 3
      colours_left
      
      if(colours_left < 2) {
        warm_colours_2 <- second_warm_colour
        
      } else {
        percent_warm <- case_when(
          colours_left == 2 ~ 40,
          colours_left == 3 ~ 33,
          colours_left == 4 ~ 25,
        )
        warm_colours_2 <- sequential(second_warm_colour,
                                     what = "value",
                                     percentage = percent_warm,
                                     plot = TRUE)
        
        warm_colours_2 <- rev(warm_colours_2[2:length(warm_colours_2)])
      }
      warm_colours <- c(warm_colours_1, warm_colours_2)
    } else {
      percent_warm <- unique(categories_with_warmer_colours$percent)
      warm_colours <- sequential(warm_starting_colour,
                                 what = "value",
                                 percentage = percent_warm,
                                 plot = TRUE)
      warm_colours <- rev(warm_colours[-1])
    }
    
    warm_colours <- remove_FF(warm_colours)
    
    categories_with_warmer_colours <- categories_with_warmer_colours %>% 
      mutate(colour = warm_colours)
    
    colours_added <- bind_rows(categories_with_cooler_colours, categories_with_warmer_colours)
    
    imported_table <- imported_table %>% 
      left_join(colours_added) %>% 
      mutate(colour = ifelse(site_series %in% c("01", "101"), zonal_colour, colour)) %>% 
      arrange(site_series)
    
    colour_scheme <- imported_table %>%
      filter(forest_non_forest == "forest") %>% 
      select(colour) %>% 
      pull
    
    df <- data.frame(
      x = imported_table %>% 
        filter(forest_non_forest == "forest") %>% 
        select(x) %>% 
        pull,
      y = as.numeric(imported_table %>% 
                       filter(forest_non_forest == "forest") %>% 
                       select(y) %>% 
                       pull),
      Site_Series = imported_table %>% 
        filter(forest_non_forest == "forest") %>% 
        select(site_series) %>% 
        pull
    )
    
    edatopic_grid <- ggplot(df, aes(x, y, width = 1, height = 1)) +
      geom_tile(aes(fill = Site_Series), colour = "grey50") + 
      xlab("Soil Nutrient Regime") +
      ylab("Soil Moisture Regime") +
      scale_fill_manual(values=unique(colour_scheme))
    
    edatopic_grid + scale_y_reverse() + scale_x_discrete(position = "top")
    
    ggsave(file.path(qml_out , paste0("edatopic_", i, ".png")))
    
    categories_and_colours <- imported_table %>% 
      mutate(classification = category) %>% 
      filter(forest_non_forest == "forest") %>% 
      select(classification, colour) %>% 
      distinct()
    
    saveRDS(categories_and_colours, file =paste0(i, "_colour.rds"))
  }
}

rd.file <- as.list(list.files(pattern = ".rds"))
rd_colours <- lapply(rd.file, function(xy) {
  my_data <- readRDS(xy)
})
unlink("*_colour.rds")

forested_colours <- do.call("rbind", rd_colours)

forest_and_non_forest_colours <- rbind(nf_colour_table, forested_colours)

distinct_colours_colour_table <- forest_and_non_forest_colours %>% 
  filter(classification %in% imported_table_full$category_mapped_for_colour) %>% 
  distinct()

distinct_colours_imported_table_full <- imported_table_full %>%
  select(category_mapped_for_colour) %>%
  distinct()

colour_duplication_check <- ifelse(
  nrow(distinct_colours_imported_table_full) != nrow(distinct_colours_colour_table),
  warning('Duplicate colours in forest_and_non_forest_colours'),
  warning('no duplicate colours present') ) 
print(colour_duplication_check)

imported_table_full <- imported_table_full %>%
  left_join(forest_and_non_forest_colours, by = c("category_mapped_for_colour" = "classification"))   

```


```{r generate .qml file, tidy = TRUE}
#map_title = "test"

# base tags that are standard to .qml files with qgis software
alpha = 255
base = newXMLNode("qgis")
addAttributes(base,
              minScale="1e+08",
              version="3.4.15-Madeira",
              mapTitle=paste0(map_title, "_v1"),
              maxScale="0",
              hasScaleBasedVisibilityFlag="0",
              styleCategories="AllStyleCategories")
flags <- newXMLNode("flags")
identifiable <- newXMLNode("Identifiable", 1)
removable <- newXMLNode("Removable", 1)
searchable <- newXMLNode("Searchable", 1)

custom_properties <- newXMLNode("customproperties")
property_a <- newXMLNode("property", attrs = c(value="false",
                                               key="WMSBackgroundLayer"))
property_b <- newXMLNode("property", attrs = c(value="false",
                                               key="WMSPublishDataSourceUrl"))
property_c <- newXMLNode("property", attrs = c(value="0",
                                               key="embeddedWidgets/count"))
property_d <- newXMLNode("property", attrs = c(value="Value",
                                               key="identify/format"))

pipe <- newXMLNode("pipe")
raster_renderer <- newXMLNode("rasterrenderer",
                              attrs = c(band="1",
                                        type="paletted",
                                        opacity="1",
                                        alphaBand="-1"))
raster_transparency <- newXMLNode("rasterTransparency")
min_max_origin <- newXMLNode("minMaxOrigin")
limits <- newXMLNode("limits", "None")
extent <- newXMLNode("extent", "WholeRaster")
stat_accuracy <- newXMLNode("statAccuracy", "Estimated")
cumulative_cut_lower <- newXMLNode("cumulativeCutLower", 0.02)
cumulative_cut_upper <- newXMLNode("cumulativeCutUpper", 0.98)
standard_deviation_factor <- newXMLNode("stdDevFactor", 2)

# colour palette which is standard for categorized colour themes in qgis, with our own colours implemented
colour_palette <- newXMLNode("colorPalette")
palette_entry <- lapply(seq_along(imported_table_full$integercode),function(val){newXMLNode("paletteEntry",
                                                                                            attrs = c(
                                                                                              label = imported_table_full$category[val],
                                                                                              value = imported_table_full$integercode[val],
                                                                                              alpha = alpha,
                                                                                              color = imported_table_full$colour[val],
                                                                                              forest_type= imported_table_full$forest_non_forest[val]))
})
addChildren(colour_palette,palette_entry)

colour_ramp <- newXMLNode("colorramp", attrs = c(type="randomcolors",
                                                 name="[source]"))

brightness_contrast <- newXMLNode("brightnesscontrast", attrs = c(brightness=0,
                                                                  contrast=0))
hue_saturation <- newXMLNode("huesaturation", attrs = c(colorizeStrength=100,
                                                        saturation=0,
                                                        colorizeRed=255,
                                                        grayscaleMode=0,
                                                        colorizeGreen=128,
                                                        colorizeBlue=128,
                                                        colorizeOn=0))

blend_mode <- newXMLNode("blendMode", 0)

addChildren(flags, list(identifiable,
                        removable,
                        searchable))
addChildren(custom_properties, list(property_a,
                                    property_b,
                                    property_c,
                                    property_d))
addChildren(pipe, list(raster_renderer,
                       brightness_contrast,
                       hue_saturation))
addChildren(raster_renderer, list(raster_transparency,
                                  min_max_origin,
                                  colour_palette,
                                  colour_ramp))
addChildren(min_max_origin, list(limits,
                                 extent,
                                 stat_accuracy,
                                 cumulative_cut_lower,
                                 cumulative_cut_upper,
                                 standard_deviation_factor))
addChildren(base, list(flags,
                       custom_properties,
                       pipe,
                       blend_mode))

writeLines(saveXML(base), file.path(qml_out, paste0(map_title, ".qml")))

if(!is.na(warning_list)) {
  print(warn_01)
}
print(colour_duplication_check)
```


```{r read geopackage file}
library(sf)

aoi <- st_read("cc_poly_example.gpkg")





```

The script below starts to solve the problem of patterning for polygons in QGIS. In the polygon_patterning folder are some example themes that have been created to represent alpine, avalanche, flood types etc. They are by no means a final draft of what the patterns should be.

```{r non_forest_qml}

get_xml_polygon <- function(type){
  return(
    getNodeSet(
      xmlParse(file = paste0("Deception_AOI/_MapUnitLegend/polygon_patterning/", type, ".qml")),
      "/document/layer")
  )
}

alpine_doc <- get_xml_polygon("alpine")
avalanche_doc <- get_xml_polygon("avalanche")
beachland_doc <- get_xml_polygon("beachland")
flood_doc <- get_xml_polygon("flood")
grassland_doc <- get_xml_polygon("grassland")
hydrogenic_doc <- get_xml_polygon("hydrogenic")
rock_doc <- get_xml_polygon("rock")
subalpine_shrub_doc <- get_xml_polygon("subalpine_shrub")
disclimax_doc <- get_xml_polygon("disclimax")
other_doc <- get_xml_polygon("other")


map_unit_match_table <- tribble(
  ~map_unit, ~match_list,
  "alpine", c("Af, Ag, Ah, Am, Wa, Az, As, At"),
  "beachland", c("Bb, Br, Bh, B"),
  "flood", c("Fl, Ff, Fa, Fm"),
  "grassland", c("Ga, Gs, Gb, Gg"),
  "hydrogenic", c("Hs, Hv, Hw, Hh"),
  "rock", c("Rc, Ro, Rt, Rl, Rd"),
  "subalpine_shrub", c("Sc, Sk"),
  "avalanche", c("Vh, Vs, Vt"),
  "disclimax", c("Xv, Xz, Xa, Xvh, Xvs, Xvt"),
)



polygons <- imported_table_full %>%
  filter(forest_non_forest == "non_forest")

polygons$match_ <- lapply(polygons$category, function(x) {
  return(
    
    case_when(
      grepl(x, map_unit_match_table$match_list[1]) ~ map_unit_match_table$map_unit[1],
      grepl(x, map_unit_match_table$match_list[2]) ~ map_unit_match_table$map_unit[2],
      grepl(x, map_unit_match_table$match_list[3]) ~ map_unit_match_table$map_unit[3],
      grepl(x, map_unit_match_table$match_list[4]) ~ map_unit_match_table$map_unit[4],
      grepl(x, map_unit_match_table$match_list[5]) ~ map_unit_match_table$map_unit[5],
      grepl(x, map_unit_match_table$match_list[6]) ~ map_unit_match_table$map_unit[6],
      grepl(x, map_unit_match_table$match_list[7]) ~ map_unit_match_table$map_unit[7],
      grepl(x, map_unit_match_table$match_list[8]) ~ map_unit_match_table$map_unit[8],
      grepl(x, map_unit_match_table$match_list[9]) ~ map_unit_match_table$map_unit[9],
    ))
})

polygons <- polygons %>% 
  mutate(match_ = ifelse(is.na(match_),
                         "other",
                         match_))

base = newXMLNode("qgis",
                  attrs = c(
                    styleCategories="AllStyleCategories",
                    simplifyDrawingTol="1",
                    simplifyDrawingHints="1",
                    simplifyLocal="1",
                    version="3.10.4-A",
                    simplifyAlgorithm="0",
                    readOnly="0",
                    hasScaleBasedVisibilityFlag="0",
                    maxScale="0",
                    labelsEnabled="0",
                    minScale="1e+08",
                    simplifyMaxScale="1",
                    mapTitle="non_forest_v1"
                  ))

flags <- newXMLNode("flags", parent = base)
identifiable <- newXMLNode("Identifiable", 1, parent = flags)
removable <- newXMLNode("Removable", 1, parent = flags)
searchable <- newXMLNode("Searchable", 1, parent = flags)

renderer_v2 <- newXMLNode("renderer-v2",
                          attrs = c(
                            forceraster="0",
                            attr="DN",
                            type="categorizedSymbol",
                            symbollevels="0",
                            enableorderby="0"
                          ),
                          parent = base)

# listy is 8 values that correlate to an example QGIS file
listy <- c(-2147483648, 0:7)


categories <- newXMLNode("categories", parent = renderer_v2)
category <- lapply(listy,function(val){
  newXMLNode(
    "category",
    attrs = c(
      symbol = val,
      value = val,
      label = val,
      render = "true"
    )
  )
})
document <- newXMLNode("document")


symbols <- newXMLNode("symbols", parent = renderer_v2)
symbol <- lapply(seq_along(listy),function(val){
  newXMLNode(
    "symbol",
    attrs = c(
      alpha = "1",
      force_rhr = "0",
      name = val,
      # value = polygons$match_[val],
      # label = polygons$category[val],
      type = "fill",
      clip_to_extent = "1"
    )
    ,
    case_when(
      listy[val] == "1" ~
        append.xmlNode(document, alpine_doc),
      listy[val] == "2" ~ append.xmlNode(document, avalanche_doc),
      listy[val] == "3" ~ append.xmlNode(document, beachland_doc),
      listy[val] == "4" ~ append.xmlNode(document, flood_doc),
      listy[val] == "5" ~ append.xmlNode(document, grassland_doc),
      listy[val] == "6" ~ append.xmlNode(document, hydrogenic_doc),
      listy[val] == "7" ~ append.xmlNode(document, rock_doc),
      listy[val] == "-2147483648" ~ append.xmlNode(document, subalpine_shrub_doc),
      listy[val] == "0" ~ append.xmlNode(document, disclimax_doc)
      # listy$thing[val] == "2" ~ append.xmlNode(document, other_doc)
    )
  )
  
})


addChildren(symbols, symbol)
addChildren(categories, category)

# base = newXMLNode("qgis_test")

writeLines(saveXML(base), file.path(qml_out, "non_forest.qml"))


```

